A sheaf--function dictionary for algebraic tori over local fields

### 2013-10-10 shanghai.tex
Broke some long lines; added a reference for l-adic sheaves; Sommes trig. |--> sommes trig.
Still looking for a reference for lisse sheaves and the etale
fundamental group . . .

We fix a prime $\ell$, invertible in $\Fq$.
We will work with constructible $\ell$-adic sheaves \cite{deligne:80a}*{\S 1.1} on $\QC$-schemes,
employing the standard formalism;
see, in particular, \cite{SGA5}*{Expos\'e VI,\S 2}, where pullback and tensor products are developed.
We also make extensive use of the external tensor product of $\ell$-adic sheaves, defined as follows: if $\mathcal{F}$ and $\mathcal{G}$ are constructible $\ell$-adic sheaves on schemes $X$ and $Y$ and $p_X : X\times Y\to X$ and $p_Y : X\times Y \to Y$ are the projections, then $\mathcal{F}\boxtimes \mathcal{G} \ceq p_X^* \mathcal{F} \otimes p_Y^*\mathcal{G}$.

@@

  To show that $L_G(f,\psi) \iso (\qcs{L}_0,\mu_0)$ we simply apply the sheaf--function dictionary
  as follows \citelist{\cite{laumon:87a}*{1.1.3} \cite{deligne:SGA4.5}*{Sommes trig.}}.
  First, recall that $L_G(f,\psi) \ceq (L(f,\psi), \mu(f,\psi))$.
  Let $k'/k$ be any degree-$n$ extension and let $\Nm : k'\to k$ be the norm map. 
  By \cite{laumon:87a}*{1.1.3.3}, 
  \[
  \trFrob{(G_{k'}\hookrightarrow G)^*L(f,\psi)} = \trFrob{\qcs{L}} \circ \Nm;
  \]
  on the other hand, 
  \[
  \trFrob{\qcs{L}} \circ \Nm = \trFrob{(G_{k'}\hookrightarrow G)^*\qcs{L}},
  \]
   by Proposition~\ref{prop:basechange}.
  %Using notation from \cite{laumon:87a}*{1.1}, this means $t_{\bullet L(f,\psi)} = t_{\bullet \qcs{L}_0}$.
  Using \cite{laumon:87a}*{Th\'eor\`eme 1.1.2}, it follows that  $L(f,\psi) \iso \qcs{L}_0$ in $\Loc(G)$.
  It is now clear that $(L(f,\psi),\mu(f,\psi)) \iso (\qcs{L}_0,\mu_0)$ in $\QCb(G)$.
  This completes the proof that $L_G : \QCf(G) \to \QC(G)$ is essentially surjective.
  Since we have already seen (in Section~\ref{sec:finite}) that $L_G$ is full and faithful, 
  it follows that $L_G$ is an equivalence.
\end{proof}

>>>>

We fix a prime $\ell$, invertible in $\Fq$.
We will work with constructible $\ell$-adic sheaves \cite{deligne:80a}*{\S 1.1}, \cite{SGA5}*{Expos\'es V, VI} 
on $\QC$-schemes, employing the standard formalism.
%see, in particular, \cite{SGA5}*{Expos\'e VI,\S 2}, where pullback and tensor products are developed.
We also make extensive use of the external tensor product of $\ell$-adic sheaves, defined as follows: if $\mathcal{F}$ and $\mathcal{G}$ are constructible $\ell$-adic sheaves on schemes $X$ and $Y$ and $p_X : X\times Y\to X$ and $p_Y : X\times Y \to Y$ are the projections, then $\mathcal{F}\boxtimes \mathcal{G} \ceq p_X^* \mathcal{F} \otimes p_Y^*\mathcal{G}$.

@@

  To show that $L_G(f,\psi) \iso (\qcs{L}_0,\mu_0)$ we simply apply the sheaf--function dictionary
  as follows \citelist{\cite{laumon:87a}*{1.1.3} \cite{deligne:SGA4.5}*{sommes trig.}}.
  First, recall that $L_G(f,\psi) \ceq (L(f,\psi), \mu(f,\psi))$.
  Let $k'/k$ be any degree-$n$ extension and let $\Nm : k'\to k$ be the norm map. 
  By \cite{laumon:87a}*{1.1.3.3}, 
  \[
  \trFrob{(G_{k'}\hookrightarrow G)^*L(f,\psi)} = \trFrob{\qcs{L}} \circ \Nm;
  \]
  on the other hand, 
  \[
  \trFrob{\qcs{L}} \circ \Nm = \trFrob{(G_{k'}\hookrightarrow G)^*\qcs{L}},
  \]
   by Proposition~\ref{prop:basechange}.
  %Using notation from \cite{laumon:87a}*{1.1}, this means $t_{\bullet L(f,\psi)} = t_{\bullet \qcs{L}_0}$.
  Using \cite{laumon:87a}*{Th\'eor\`eme 1.1.2}, it follows that  $L(f,\psi) \iso \qcs{L}_0$ in $\Loc(G)$.
  It is now clear that $(L(f,\psi),\mu(f,\psi)) \iso (\qcs{L}_0,\mu_0)$ in $\QCb(G)$.
  This completes the proof that $L_G : \QCf(G) \to \QC(G)$ is essentially surjective.
  Since we have already seen (in Section~\ref{sec:finite}) that $L_G$ is full and faithful, 
  it follows that $L_G$ is an equivalence.
\end{proof}
